---
title: "Prepping for dada2"
author: "Vicki Hertzberg"
date: "2/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

To work along with the dada2 lecture on 3/3, you will need to install a number of programs. Most of these come from Bioconductor.

You will need the following programs from CRAN:

ggplot2

You will need to install the following programs from Bioconductor:

dada2
phyloseq
ShortRead
DECIPHER

To install a program from Bioconductor, you will first need to install Bioconductor. To do this, execute the following code IN THE CONSOLE WINDOW:



if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.12")

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.12")
```




Next install the other programs. For instance to install dada2, use the following code:

```{r}

#Remove comment symbol from next line for it to work. I am keeping the next line commented, however, because I do not need to install dada2, it is already installed.

#BiocManager::install("dada2")

```

To install the other programs, use a similar chunk of code, this time substituting the package name for dada2.

Next check on the package version of all of the programs you have installed:

```{r}

packageVersion("ggplot2")
packageVersion("dada2")
packageVersion("DECIPHER")
packageVersion("ShortRead")
packageVersion("ggplot2")

```

There are some other files you will need to download.

First you wil need to download the mothur MiSeq SOP dataset that you can find at http://mothur.s3.us-east-2.amazonaws.com/wiki/miseqsopdata.zip

Unzip this download and store the folder on your hard drive (not your desktop), and remember the path to it. You should have 45 files. 

Also, go to https://zenodo.org/record/3986799#.YC7mEi2cbUI and downlod the silva_nr99_v138_train_set.fa.gz, the silva_nr99_v138_wSpecies_train_set.fa.gz, and the silva_species_assignment_v138.fa.gz files. 

Store these files in the same place where you are storing the mothur example dataset. DO NOT UNZIP.

We are now ready to take on dada2 on 3/3/21.



